A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-11 main.tex
[WIP] Resolve conflict
Remove itemized references.

% Typing rules for a multi-stage calculus
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp\ are rules for a multi-stage calculus.
They are corresponding to brackets, escaping a code, making a stage abstraction, 
application of $\epsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

% \TConv
The main technical point of this paper is \TConv\ and type equivalence rules needed by \TConv.
This rule allows ut to replace a type with another type that is equivalent.
In a type system which includes dependent types, this kind of rule is essential
because two types which have different shapes may be equivalent.
For example, when we use a matrix type which have their sizes ($\textrm{Mat}\ n\ m$),
$\textrm{Mat}\ 5\ 3$ is equivalent to $\textrm{Mat}\ (4+1)\ (1+2)$ obviously.

@@

% 2つの設計手法と今回採用した理由
When we design type equivalence rules of \LMD, there are two design choices.
One is defining type equivalence by $\beta$-equality after we define $\beta$-reduction of types.
Another is defining type equivalence directly by a combination of type equivalence rules.
We adopt the latter one because it is convenient to handle CSP in the type equivalence.

@@

\begin{figure}
	\begin{center}
		\infrule{\G\V \tau \E \sigma :: *@A \andalso \G,x:\tau @A \V M \E N : \rho @A}{\G\V\lambda x:\tau.M \E \lambda x:\sigma.N : (\Pi x:\tau.\rho)@A}{\QAbs} \\[2mm]
		\infrule{\G\V M \E L : (\Pi x:\sigma.\tau)@A \andalso \G\V N \E O : \sigma@A}{\G\V M\ N \E L\ O : \tau[x \mapsto N]@A}{\QApp} \\[2mm]
		\infrule{\G\V M \E N : \tau @{A\alpha}}{\G\V \TB_\alpha M \E \TB_\alpha N : \TW_\alpha \tau @A}{\QTB} \andalso
		\infrule{\G\V M \E N : \TW_\alpha \tau @A}{\G\V \TBL_\alpha M \E \TBL_\alpha N : \tau @{A\alpha}}{\QTBL} \\[2mm]
		\infrule{\G\V M\E N : \tau @A \andalso \alpha \notin \FTV(\G)\cup\FTV(A)}{\G\V \Lambda\alpha.M \E \Lambda\alpha.N : \forall\alpha.\tau @A}{\QGen} \\[2mm]
		\infrule{\G\V M \E N:\forall\alpha.\tau @A}{\G\V M\ \epsilon \E N\ \epsilon : \tau[\alpha \mapsto \epsilon]@A}{\QIns}\andalso
		\infrule{\G\V M \E N : \tau @A}{\G\V\%_\alpha M \E \%_\alpha N : \tau @{A\alpha}}{\QCsp} \\[2mm]
		\infrule{\G\V M:\tau @A}{\G\V M\E M : \tau @A}{\QRefl} \andalso
		\infrule{\G\V M\E N : \tau @A}{\G\V N\E M : \tau @A}{\QSym} \\[2mm]
		\infrule{\G\V M\E N : \tau @A \andalso \G\V N\E L : \tau @A}{\G\V M\E L : \tau @A}{\QTrans} \\[2mm]
		\infrule{\G,x:\sigma@A\V M:\tau @A \andalso \G\V N:\sigma@A}{\G\V(\lambda x:\sigma.M)\ N\E M[x\mapsto N] : \tau[x \mapsto N]@A}{\QBeta} \\[2mm]
		% \infrule{\G\V M:(\Pi x:\sigma.\tau)@A \andalso x\notin \text{FV}(M)}{\G\V(\lambda x:\sigma.M\ x)\E M: (\Pi x:\sigma.\tau)@A}{\QEta} \\[2mm]
		\infrule{\G\V M \E N : \tau @A}{\G\V \TBL_\alpha(\TB_\alpha M) \E N : \tau @A}{\QTBLTB} \\[2mm]
		\infrule{\G\V (\Lambda\alpha.M) : \forall\alpha.\tau @A}{\G\V (\Lambda\alpha.M)\ \epsilon \E M[\alpha \mapsto \epsilon] : \tau[\alpha \mapsto \epsilon]@A}{\QLambda} \\[2mm]
		\infrule{\G\V M:\tau @{A\alpha} \andalso \G\V M:\tau @A}{\G\V\%_\alpha M \E M : \tau @{A\alpha}}{\QPercent}
		\caption{Term Equivalence Rules}
		\label{fig:term-equivalence-rules}
	\end{center}
\end{figure}

@@

\begin{enumerate}
	\item Done: On Cross-Stage Persistence in Multi-Stage\cite{Hanada2014}

	      CSPも入りました
	\item Done: Eliminating Array Bound Checking Through Dependent Types\cite{Xi98}
	\item Done: MetaML and Multi-stage Programming with Explicit Annotations\cite{MetaML}
	\item Done: Idris, a general-purpose dependently typed programming language: Design and implementation\cite{brady2013idris}
	\item Done: A Logical Foundation for Environment Classifiers\cite{Tsukada}
	\item Done: Environment classifiers\cite{taha2003environment}
	\item Done: A framework for defining logics\cite{harper1993framework}

	      $\Sigma$ の使い方を確認した
	\item Done: The Design and Implementation of {BER} MetaOCaml - System Description\cite{oleg2014}
	\item Done: Dependent types in practical programming\cite{xi1999dependent}

	      Section of Applicationを読んだ。Dead Code EliminationやLoop Unrollingにも使えるらしい。

	\item Done: A dependently typed assembly language\cite{xi2001dependently}

	      DTALの定義と制約solverを用いた型検査の定義
	\item Done: Refined Environment Classifiers\cite{kiselyov2016refined}
	\item Done: Staging with control: type-safe multi-stage programming with control operators\cite{oishi2017staging}
	\item Done: Logical Modalities and Multi-Stage Programming\cite{benaissa1999logical}
	      % \item \red{Partial evaluation and automatic program generation}\cite{jones1993partial}
	      % \item \red{Efficient multi-level generating extensions for program specialization}\cite{gluck1995efficient}
	      % \item \red{Run-time code generation and Modal-ML}\cite{wickline1998run}
	      % \item \red{C and tcc: a language and compiler for dynamic code generation}\cite{poletto1999c}
	      % \item \red{Run-time bytecode specialization}\cite{masuhara2001run}
	      % \item \red{A tour of Tempo: A program specializer for the C language}\cite{consel2004tour}
	      % \item \red{Optimizing ML with run-time code generation}\cite{lee1996optimizing}
	      % \item \red{Efficient incremental run-time specialization for free}\cite{marlet1999efficient}
\end{enumerate}

@@

>>>>

% Typing rules for a multi-stage calculus
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp\ are rules for a multi-stage calculus.
They are corresponding to brackets, escape, making a stage abstraction, 
application of $\epsilon$ stage to a stage abstraction, and cross-stage persistence, respectively.
Please check Hanada and Igarashi \cite{Hanada2014} and Tsukada and Igarashi \cite{Tsukada} for details of these rules.

% \TConv
The main technical point of this paper is \TConv\ and type equivalence rules needed by \TConv.
This rule allows us to replace a type with another type that is equivalent.
In a type system which includes dependent types, this kind of rule is essential
because two types which have different shapes may be equivalent.
For example, when we use a vector type with its size ($\textrm{Vector}\ n$),
$\textrm{Vector}\ 5$ is equivalent to $\textrm{Vector}\ (4+1)$ obviously although they are not equivalent in apparently.

@@

@@

\begin{figure}
	\begin{center}
		\infrule{\G\V \tau \E \sigma :: *@A \andalso \G,x:\tau @A \V M \E N : \rho @A}{\G\V\lambda x:\tau.M \E \lambda x:\sigma.N : (\Pi x:\tau.\rho)@A}{\QAbs} \\[2mm]
		\infrule{\G\V M \E L : (\Pi x:\sigma.\tau)@A \andalso \G\V N \E O : \sigma@A}{\G\V M\ N \E L\ O : \tau[x \mapsto N]@A}{\QApp} \\[2mm]
		\infrule{\G\V M \E N : \tau @{A\alpha}}{\G\V \TB_\alpha M \E \TB_\alpha N : \TW_\alpha \tau @A}{\QTB} \andalso
		\infrule{\G\V M \E N : \TW_\alpha \tau @A}{\G\V \TBL_\alpha M \E \TBL_\alpha N : \tau @{A\alpha}}{\QTBL} \\[2mm]
		\infrule{\G\V M\E N : \tau @A \andalso \alpha \notin \FTV(\G)\cup\FTV(A)}{\G\V \Lambda\alpha.M \E \Lambda\alpha.N : \forall\alpha.\tau @A}{\QGen} \\[2mm]
		\infrule{\G\V M \E N:\forall\alpha.\tau @A}{\G\V M\ \epsilon \E N\ \epsilon : \tau[\alpha \mapsto \epsilon]@A}{\QIns}\andalso
		\infrule{\G\V M \E N : \tau @A}{\G\V\%_\alpha M \E \%_\alpha N : \tau @{A\alpha}}{\QCsp} \\[2mm]
		\infrule{\G\V M:\tau @A}{\G\V M\E M : \tau @A}{\QRefl} \andalso
		\infrule{\G\V M\E N : \tau @A}{\G\V N\E M : \tau @A}{\QSym} \\[2mm]
		\infrule{\G\V M\E N : \tau @A \andalso \G\V N\E L : \tau @A}{\G\V M\E L : \tau @A}{\QTrans} \\[2mm]
		\infrule{\G,x:\sigma@A\V M:\tau @A \andalso \G\V N:\sigma@A}{\G\V(\lambda x:\sigma.M)\ N\E M[x\mapsto N] : \tau[x \mapsto N]@A}{\QBeta} \\[2mm]
		% \infrule{\G\V M:(\Pi x:\sigma.\tau)@A \andalso x\notin \text{FV}(M)}{\G\V(\lambda x:\sigma.M\ x)\E M: (\Pi x:\sigma.\tau)@A}{\QEta} \\[2mm]
		\infrule{\G\V (\Lambda\alpha.M) : \forall\alpha.\tau @A}{\G\V (\Lambda\alpha.M)\ \epsilon \E M[\alpha \mapsto \epsilon] : \tau[\alpha \mapsto \epsilon]@A}{\QLambda} \\[2mm]
		\infrule{\G\V M \E N : \tau @A}{\G\V \TBL_\alpha(\TB_\alpha M) \E N : \tau @A}{\QTBLTB} \andalso
		\infrule{\G\V M:\tau @{A\alpha} \andalso \G\V M:\tau @A}{\G\V\%_\alpha M \E M : \tau @{A\alpha}}{\QPercent}
		\caption{Term Equivalence Rules}
		\label{fig:term-equivalence-rules}
	\end{center}
\end{figure}

@@

% \begin{enumerate}
% 	\item Done: On Cross-Stage Persistence in Multi-Stage\cite{Hanada2014}

% 	      CSPも入りました
% 	\item Done: Eliminating Array Bound Checking Through Dependent Types\cite{Xi98}
% 	\item Done: MetaML and Multi-stage Programming with Explicit Annotations\cite{MetaML}
% 	\item Done: Idris, a general-purpose dependently typed programming language: Design and implementation\cite{brady2013idris}
% 	\item Done: A Logical Foundation for Environment Classifiers\cite{Tsukada}
% 	\item Done: Environment classifiers\cite{taha2003environment}
% 	\item Done: A framework for defining logics\cite{harper1993framework}

% 	      $\Sigma$ の使い方を確認した
% 	\item Done: The Design and Implementation of {BER} MetaOCaml - System Description\cite{oleg2014}
% 	\item Done: Dependent types in practical programming\cite{xi1999dependent}

% 	      Section of Applicationを読んだ。Dead Code EliminationやLoop Unrollingにも使えるらしい。

% 	\item Done: A dependently typed assembly language\cite{xi2001dependently}

% 	      DTALの定義と制約solverを用いた型検査の定義
% 	\item Done: Refined Environment Classifiers\cite{kiselyov2016refined}
% 	\item Done: Staging with control: type-safe multi-stage programming with control operators\cite{oishi2017staging}
% 	\item Done: Logical Modalities and Multi-Stage Programming\cite{benaissa1999logical}
% 	\item \red{Partial evaluation and automatic program generation}\cite{jones1993partial}
% 	\item \red{Efficient multi-level generating extensions for program specialization}\cite{gluck1995efficient}
% 	\item \red{Run-time code generation and Modal-ML}\cite{wickline1998run}
% 	\item \red{C and tcc: a language and compiler for dynamic code generation}\cite{poletto1999c}
% 	\item \red{Run-time bytecode specialization}\cite{masuhara2001run}
% 	\item \red{A tour of Tempo: A program specializer for the C language}\cite{consel2004tour}
% 	\item \red{Optimizing ML with run-time code generation}\cite{lee1996optimizing}
% 	\item \red{Efficient incremental run-time specialization for free}\cite{marlet1999efficient}
% \end{enumerate}

@@

% 2つの設計手法と今回採用した理由
% Conclusion送りか?
% When we design type equivalence rules of \LMD, there are two design choices.
% One is defining type equivalence by $\beta$-equality after we define $\beta$-reduction of types.
% Another is defining type equivalence directly by a combination of type equivalence rules.
% We adopt the latter one because it is convenient to handle CSP in the type equivalence.
